---
title: "Conditional model for Genome-Wide Selection"
subtitle: ""
author: "Moi Exposito-Alonso"
date: '`r Sys.Date()`'
output:
  # pdf_document:
  # header-inclusdes:
  #  - \usepackage{amsmath}
  #  - \usepackage{caption}
  #  - \usepackage{subcaption}
  #  - \usepackage{graphicx}
  html_document: default
  header-includes:
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
set.seed(1)
```


```{r packages, include=FALSE}
## Load packages
library(devtools)
library(dplyr)
library(ggplot2)
library(cowplot)
library(xtable)

load_all('.')
# library('gws')
library(moiR)
library(bigsnpr)
library(bigstatsr)

```

*****
## 1 Subset random positions 
```{r}
### Real data
# genomes<-readRDS("genome.rda")
data(genomes)
G <- genomes$genotypes

X.random = inputena.mat(G[,sample(1:ncol(G),size = 600)])
N=nrow(X.random)
M=ncol(X.random)

qplot(fn(X.random), xlab="allele dossage",main="")

```

## 2 Subset a window
```{r}

X.window =inputena.mat(G[,c(1:600)]) # Instead of sampling, I get them consecutive
N=nrow(X.window)
M=ncol(X.window)

qplot(fn(X.window), xlab="allele dossage",main="")

```

<!-- ###################################################################### -->

## 3 Association with a Gaussian trait generated from different architectures
### Gaussian effects
```{r}

eff.gauss=rnorm(M,mean = 0,sd = 1)
qplot(eff.gauss,geom="histogram",main="Simulated weak Gaussian SNP effects")

Y.random.gauss=meanvarcent(X.random %*% eff.gauss)
Y.window.gauss=meanvarcent(X.window %*% eff.gauss)

```

### Discrete large effects
```{r}
prob=0.005
eff.skewed=sample(c(-1,0,1),replace = TRUE, prob = c(prob,1-2*prob,prob),size = M)
# eff.skewed=rexp(n=M,rate = 10e6)

qplot(eff.skewed,geom="histogram",main="Simulated discrete large SNP effects")

Y.random.skewed=meanvarcent(X.random %*% eff.skewed)
Y.window.skewed=meanvarcent(X.window %*% eff.skewed)
Y.window.skewed2=meanvarcent((X.window %*% eff.skewed ) *(X.window %*% eff.skewed ))
Y.window.skewed2=meanvarcent((X.window %*% eff.skewed ) *(X.window %*% eff.skewed ))

plot(Y.window.skewed,Y.window.skewed2)


Y.window.mixture= meanvarcent((X.window %*% eff.skewed ) + (X.window %*% (eff.gauss/10) ))
Y.window.mixture= meanvarcent((X.window %*% eff.skewed ) * (X.window %*% (eff.gauss/10) ))


X.window %*% eff.skewed


```

## 4 Calculate coefficients
```{r, fig.asp = 0.5}


b.random.gauss=cgwa(X.random,Y.random.gauss) 
b.random.skewed=cgwa(X.random,Y.random.skewed) 
b.window.gauss=cgwa(X.window,Y.window.gauss) 
b.window.skewed=cgwa(X.window,Y.window.skewed) 

bgwa.random.gauss=gwa(X.random,Y.random.gauss) 
bgwa.random.skewed=gwa(X.random,Y.random.skewed) 
bgwa.window.gauss=gwa(X.window,Y.random.gauss) 
bgwa.window.skewed=gwa(X.window,Y.random.skewed) 

blasso.random.gauss=lassogwa(X.random,Y.random.gauss) 
blasso.random.skewed=lassogwa(X.random,Y.random.skewed) 
blasso.window.gauss=lassogwa(X.window,Y.window.gauss) 
blasso.window.skewed=lassogwa(X.window,Y.window.skewed) 

bridge.random.gauss=ridgegwa(X.random,Y.random.gauss) 
bridge.random.skewed=ridgegwa(X.random,Y.random.skewed) 
bridge.window.gauss=ridgegwa(X.window,Y.window.gauss) 
bridge.window.skewed=ridgegwa(X.window,Y.window.skewed) 


plot(ridgegwa(X.random,Y.random.skewed) ,eff.skewed)
plot(ridgegwa(X.random,Y.random.skewed,type = "ld") ,eff.skewed,lambda=100)
plot( 
     ridgegwa(X.random,Y.random.skewed,type = "penalized",k = 100,x0 = 0.99,lambda = 1)
  ,eff.skewed)

plot( 
     ridgegwa(X.window,Y.window.skewed,type = "penalized",k = 100,x0 = 0.9,lambda = 1)
  ,eff.skewed)

plot( 
     ridgegwa(X.window,Y.window.gauss,type = "penalized",k = 10,x0 = 0.5,lambda = 1)
  ,eff.gauss)
plot( 
     ridgegwa(X.random,Y.random.gauss,type = "penalized",k=1,x0 = 0.5,lambda=.1)
  ,eff.gauss)


plot( 
     # ridgegwa(X.window,Y.window.skewed,type = "ldpenalized")
     ridgegwa(X.window,Y.window.skewed,type = "penalized")
  ,eff.skewed)
plot( 
      ridgegwa(X.window,Y.window.gauss,type = "penalized")
     # ridgegwa(X.window,Y.window.gauss,type = "ldpenalized")
  ,eff.gauss)

plot( 
      # ridgegwa(X.random,Y.random.gauss,type = "penalized")
     ridgegwa(X.random,Y.random.gauss,type = "ldpenalized")
  ,eff.gauss)

plot( 
     ridgegwa(X.random,Y.random.gauss,type = "penalized")
    ,gwa(X.random,Y.random.gauss)
)
plot( 
     ridgegwa(X.random,Y.random.gauss,type = "penalized")
    ,gwa(X.random,Y.random.gauss)
)

plot( 
     ridgegwa(X.window,Y.window.gauss,type = "penalized")
    ,gwa(X.window,Y.window.gauss)
)


plot( 
      ridgegwa(X.window,Y.window.mixture,type = "penalized")
  ,eff.skewed*eff.gauss/10)


cgwa(X.random,Y.random.gauss) %>% hist
cgwa(X.random,Y.random.skewed) %>% hist
cgwa(X.window,Y.window.gauss) %>% hist
cgwa(X.window,Y.window.skewed) %>% hist

plot(cgwa(X.random,Y.random.skewed) ,eff.skewed)

warning(FALSE)
bridge.random.gauss=ridgegwa(X.random,Y.random.gauss,lambda = 1) 
cor.test(eff.gauss,bridge.random.gauss)$estimate
bridge.random.gauss=ridgegwa(X.random,Y.random.gauss,lambda = 0.1) 
cor.test(eff.gauss,bridge.random.gauss)$estimate
bridge.random.gauss=ridgegwa(X.random,Y.random.gauss,lambda = .01) 
cor.test(eff.gauss,bridge.random.gauss)$estimate
bridge.random.gauss=ridgegwa(X.random,Y.random.gauss,lambda = .001) 
cor.test(eff.gauss,bridge.random.gauss)$estimate
bridge.random.gauss=ridgegwa(X.random,Y.random.gauss,lambda = .0001) 
cor.test(eff.gauss,bridge.random.gauss)$estimate

bridge.random.gauss=regularridgegwa(X.random,Y.random.gauss,lambda = .0001) 
bridge.random.gauss=regularridgegwa(X.random,Y.random.gauss,lambda = .001) 
cor.test(eff.gauss,bridge.random.gauss) 

lm(eff.gauss~bridge.random.gauss) %>% summary
bridge.random.gauss=cgwa(X.random,Y.random.gauss) 
lm(eff.gauss~bridge.random.gauss) %>% summary


# sum(abs(eff.gauss))
# sum(abs(bgwa.random.gauss))
# sum(abs(b.random.gauss))
# 
# sum(abs(eff.skewed))
# sum(abs(bgwa.random.skewed))
# sum(abs(b.random.skewed))
# 
# 
# sum(abs(eff.gauss))
# sum(abs(bgwa.window.gauss))
# sum(abs(b.window.gauss))
# 
# sum(abs(eff.skewed))
# sum(abs(bgwa.window.skewed))
# sum(abs(b.window.skewed))


```

## 4 Visualize differences in \beta estimates
#### Simulation with SNPs distributed randomly in the genome (low LD)
The plot below indicates that the best performing for a sparse genetic architecture is Lasso, while for a polygenic architecture, with many Gaussian effects, the conditional model is still the best.

```{r,fig.width=6, fig.asp=1.5}

grid1<-plot_grid(ncol=2,
addggregression(ggdotscolor(x=eff.skewed,y=bgwa.random.skewed) + ggtitle('marginal | skewed')) ,
addggregression(ggdotscolor(x=eff.gauss,y=bgwa.random.gauss) + ggtitle('marginal | gauss')) ,

addggregression(ggdotscolor(x=eff.skewed,y=b.random.skewed) + ggtitle('conditional | skewed')) ,
addggregression(ggdotscolor(x=eff.gauss,y=b.random.gauss) + ggtitle('conditional | gauss')) ,

addggregression(ggdotscolor(x=eff.skewed,y=blasso.random.skewed) + ggtitle('lasso | skewed')) , 
addggregression(ggdotscolor(x=eff.gauss,y=blasso.random.gauss) + ggtitle('lasso | gauss')) ,

addggregression(ggdotscolor(x=eff.skewed,y=bridge.random.skewed) + ggtitle('ridge | skewed')) , 
addggregression(ggdotscolor(x=eff.gauss,y=bridge.random.gauss) + ggtitle('ridge | gauss')) 
) + ggtitle('Random SNPs')

grid1
save_plot(file='figs/random_snps_simulationgwa.pdf', grid1,base_width = 10,base_height = 12)






```

#### Simulation with SNPs distributed contiguously (high LD)

The plot below indicates that once LD is high, only the conditional model performs relatively well. The marginal and the lasso perform poorly.

```{r,fig.width=6, fig.asp=1.5}

grid2<-plot_grid(ncol=2,
addggregression(ggdotscolor(x=eff.skewed,y=bgwa.window.skewed) + ggtitle('marginal | skewed')) ,
addggregression(ggdotscolor(x=eff.gauss,y=bgwa.window.gauss) + ggtitle('marginal | gauss')) ,

addggregression(ggdotscolor(x=eff.skewed,y=b.window.skewed) + ggtitle('conditional | skewed')) ,
addggregression(ggdotscolor(x=eff.gauss,y=b.window.gauss) + ggtitle('conditional | gauss')) ,

addggregression(ggdotscolor(x=eff.skewed,y=blasso.window.skewed) + ggtitle('lasso | skewed')) , 
addggregression(ggdotscolor(x=eff.gauss,y=blasso.window.gauss) + ggtitle('lasso | gauss')) ,

addggregression(ggdotscolor(x=eff.skewed,y=bridge.window.skewed) + ggtitle('bridge | skewed')) , 
addggregression(ggdotscolor(x=eff.gauss,y=bridge.window.gauss) + ggtitle('bridge | gauss')) 

)
grid2
save_plot(file='figs/window_snps_simulationgwa.pdf', grid2,base_width = 10,base_height = 12)

```

<!-- ```{r} -->

<!-- d=data.frame(eff=eff.skewed,b,bgwa) -->
<!-- d$position=1:nrow(d) -->

<!-- ggplot(d) + geom_line(aes(y=abs(bgwa),x=position),color="black",alpha=0.5) + geom_line(aes(y=abs(b),x=position),color="red",alpha=0.5) +  ylab(expression(paste("Absolute ",beta))) -->

<!-- ggplot(d) + geom_line(aes(y=abs(bgwa),x=position),color="black",alpha=0.5) + geom_line(aes(y=abs(b),x=position),color="red",alpha=0.5) +  ylab(expression(paste("Absolute ",beta))) + -->
<!--   xlim(c(0,100)) -->

<!-- ggplot(d) + geom_line(aes(y=abs(bgwa) - abs(b),x=position),color="black",alpha=0.5)+ -->
<!--   geom_hline(yintercept = 0,lty="dashed") + -->
<!--   ylab(expression(paste(beta[gwa]," - ",beta," (absolute difference)"))) + -->
<!--   xlim(c(0,100)) -->

<!-- ``` -->

