---
title: "Genome-wide LD change as a signature of selection in experimental populations"
# subtitle: "EQUATIONS"
author: "Moi Exposito-Alonso"
# date: '`r Sys.Date()`'
output:
  pdf_document:
  header-inclusdes:
   - \usepackage{amsmath}
   - \usepackage{caption}
   - \usepackage{subcaption}
   - \usepackage{graphicx}
  #html_document: default
  #header-includes:
  #- \setlength{\parindent}{4em}
  #- \setlength{\parskip}{0em}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
set.seed(1)
```


```{r packages, include=FALSE}
## Load packages
library(devtools)
library(dplyr)
library(ggplot2)
library(cowplot)
library(latex2exp)

devtools::load_all('.')
# library('gws')
library(moiR)

```
 


\
\begin{center}
\begin{tabular}{ c | c  c}
     &       B            &     b       \\ 
  \hline
  A & $(1+s_A)(1+s_B) $&$ (1+s_{A})  $\\  
  a & $(1+ s_{B})$&$ 1$    
\end{tabular}
\end{center}
\

<!-- \ -->
<!-- \begin{center} -->
<!-- \begin{tabular}{ c c c c c c c c c} -->
<!--      C & & & &     c       \\  -->
<!--   \hline -->
<!--      B && b && B && b\\  -->
<!--   \hline -->
<!--      A & a & A & a & A & a & A & a\\  -->
<!--   \hline -->
<!--   $(1+s_A)(1+s_B) $&$ (1+s_{A}) (1-s_{B})  $\\   -->
<!--   $(1- s_{A})(1+ s_{B})$&$ (1-s_A)(1-s_B)          $     -->
<!-- \end{tabular} -->
<!-- \end{center} -->
<!-- \ -->


## The D and R do not have an 1 to 1 relationship
```{r}

x1<-seq(0.01,0.99,0.1)
x2<-seq(0.01,0.99,0.1)
x3<-seq(0.01,0.99,0.1)
x4<-seq(0.01,0.99,0.1)

allxs<-expand.grid(x1,x2,x3,x4)

ggdotscolor(getD(allxs),getR(allxs), xlab='D',ylab="Felsenstein's R") + scale_y_log10()

```

# LD (r2, D nor R) does not change with multiplicative selection without epistasis
```{r}
s<-seq(-0.5,+0.5,0.05)
t<-seq(-0.5,+0.5,0.05)
E<-0

ts<-expand.grid(t,s,E)

w1<-(1+ts[,1]+ts[,2])
w1<-(1+ts[,1])*(1+ts[,2])
w1<-(1+ts[,1])*(1+ts[,2]) *(1+ts[,3])
w2<-(1+ts[,2])
w3<-(1+ts[,1])
w4<-rep(1,nrow(ts))
allw<-cbind(w1,w2,w3,w4)

qplot(x=fn(getR2change(allw)), xlab=TeX('r^2 change'))
qplot(x=fn(getDchange(allw)), xlab=TeX('D change'))
qplot(x=fn(getRchange(allw)), xlab=TeX('R change'))


```

# LD does change with multiplicative selection and multiplicative epistasis
```{r}
s<-seq(-0.5,+0.5,0.05)
t<-seq(-0.5,+0.5,0.05)
E<-seq(-0.5,+0.5,0.05)

ts<-expand.grid(t,s,E)

w1<-(1+ts[,1]+ts[,2])
w1<-(1+ts[,1])*(1+ts[,2])
w1<-(1+ts[,1])*(1+ts[,2]) *(1+ts[,3])
w2<-(1+ts[,2])
w3<-(1+ts[,1])
w4<-rep(1,nrow(ts))
allw<-cbind(w1,w2,w3,w4)


qplot(x=fn(getR2change(allw)), xlab=TeX('r^2 change'))
qplot(x=fn(getDchange(allw)), xlab=TeX('D change'))
qplot(x=fn(getRchange(allw)), xlab=TeX('R change'))

```


# LD does change with multiplicative selection and power epistasis
```{r}
s<-seq(-0.5,+0.5,0.05)
t<-seq(-0.5,+0.5,0.05)
E<-seq(-0.5,+0.5,0.05)

ts<-expand.grid(t,s,E)

w1<-((1+ts[,1])*(1+ts[,2]))^(1+ts[,3])
w2<-(1+ts[,2])
w3<-(1+ts[,1])
w4<-rep(1,nrow(ts))
allw<-cbind(w1,w2,w3,w4)

qplot(x=fn(getR2change(allw)), xlab=TeX('r^2 change'))
qplot(x=fn(getDchange(allw)), xlab=TeX('D change'))
qplot(x=fn(getRchange(allw)), xlab=TeX('R change'))

```


# Does LD change with a 3rd locus unders selection?

```{r}


Dc<-0.02
pc<-0.1
sc<-0.8

s<-seq(-0.5,+0.5,0.05)
t<-seq(-0.5,+0.5,0.05)
E<-seq(-0.5,+0.5,0.05)
E<-0

ts<-expand.grid(t,s,E)

w1<-(((1+ts[,1])*(1+ts[,2]))^(1+ts[,3]) )  [1]
w2<-(1+ts[,2])                           [1]
w3<-(1+ts[,1])                           [1]   
w4<-rep(1,nrow(ts))                      [1]   
allw<-cbind(w1,w2,w3,w4)

a=0.2
b=0.3
c=0.6

Dab<-0.05
Dbc<-0.03
Dac<-0.02
  

freq3<-function(G='ABC',a,b,c,Dab,Dac,Dbc){
  
  a*b*c  * (1+Dab+Dac+Dbc - (Dab*Dac)-(Dab*Dbc)-(Dac*Dbc) +(Dab*Dac*Dbc))

}
ABC<- a*b*c  + (Dab+Dac+Dbc - (Dab*Dac)-(Dab*Dbc)-(Dac*Dbc) +(Dab*Dac*Dbc))
aBC<- (1-a)*b*c + (-Dab+Dac+Dbc - (Dab*Dac)-(Dab*Dbc)-(Dac*Dbc) +(Dab*Dac*Dbc))
AbC<- a*(1-b)*c * (1+Dab)
abC<- (1-a)*(1-b)*c * (1+Dab)
ABc<- a*b*(1-c) * (1+Dab)
aBc<- (1-a)*b*(1-c) * (1+Dab)
Abc<- a*(1-b)*(1-c) * (1+Dab)
abc<- (1-a)*(1-b)*(1-c) * (1+Dab)


x1C =0.4
x4C =0.4
x2C =0.1
x3C =0.1

e=0

w1<-((1+sa)*(1+sb))^(1+e) 
w2<-(1+sb)                           
w3<-(1+sa)                           
w4<-1

sc=0.8

w1<- (1-x1C) * w1 + x1C *(w1*(1+sc))
w4<- (1-x4C) * w4 + x4C *(w4*(1+sc))
w2<- (1-x2C) * w2 + x2C *(w2*(1+sc))
w3<- (1-x3C) * w3 + x3C *(w3*(1+sc))

(w1*w4) / (w2*w3)


# x1p = (x1*pc +Dc) * (w1+sc) + (1-(x1*pc+Dc)) *(w1)
# x2p = (x2*pc -Dc) * (w2+sc) + (1-(x2*pc-Dc)) *(w2)
# x3p = (x3*pc -Dc) * (w3+sc) + (1-(x3*pc-Dc)) *(w3)
# x4p = (x4*pc +Dc) * (w4+sc) + (1-(x4*pc+Dc)) *(w4)
# 
# x1p= ()
# 
# w_hat=(x1p+x2p+x3p+x4p)
# x1p=x1p/w_hat
# x2p=x2p/w_hat
# x3p=x3p/w_hat
# x4p=x4p/w_hat
# 
# pp<-x1p+x2p
# qp<-x3p+x4p
# pp+qp
# p=x1+x2
# q=x3+x4
# 
# ((x1p*x4p) / (x2p*x3p) ) - (x1*x4) / (x2*x3) 
# ((x1p*x4p) - (x2p*x3p) ) - (x1*x4) - (x2*x3) 
# 
# (((x1p*x4p) - (x2p*x3p) )/sqrt(pp*qp*(1-pp)*(1-qp)) ) - ( ((x1*x4) - (x2*x3)) /sqrt(p*q*(1-p)*(1-q)) )
# 
# 
# x

```



```{r,fig.asp=1}

X<-get_topSNPs(100,navalue = 0.5)
dim(X)
X[,1] <-X[,1] * (-1)

nsnps=3

ldfocal<-c()
Dwith3rd<-c()

for(i in 1:1000){
  Xs<-X[,sample(1:100,nsnps)]
  Xs[Xs==(-1)]<-0
  
  s<-rep(0.5,nsnps)
  # e<- 0.8
  e<- 0
  
  haps<-apply(Xs,1,function(i) prod(i))
  w<- apply(1+ s*Xs, 1, function(i){prod(i)})
  w<-w^(1+(e*haps))
  
  ldnow_d<-ldCnow(Xs,R = FALSE,dolog = FALSE, addone=FALSE) %>% upperTmat()
  ldnow<-ldCnow(Xs,R = TRUE,dolog = FALSE, addone=FALSE) %>% upperTmat()
  ldnext<-ldCnext(Xs,w,R=TRUE,dolog = FALSE,addone=FALSE) %>% upperTmat()
  ldnext/ldnow
  Dwith3rd<-append(Dwith3rd,mean(ldnow_d[-1],na.rm=TRUE))
  ldfocal<-append(ldfocal,mean((ldnext/ldnow)[1],na.rm=TRUE))
  
}

qplot(ldfocal,xlab='R between two focal SNPs',xlim=c(0,2))+
  geom_vline(xintercept = mean(ldfocal,na.rm = T), col='firebrick', lty='dashed',lwd=2)
# hist(ldfocal)
mean(ldfocal,na.rm = T)



for(i in 1:1000){
  Xs<-X[,sample(1:100,nsnps)]
  Xs[Xs==(-1)]<-0
  
  s<-rep(0.5,nsnps)
  # e<- 0.8
  e<- 1
  
  haps<-apply(Xs,1,function(i) prod(i))
  w<- apply(1+ s*Xs, 1, function(i){prod(i)})
  w<-w^(1+(e*haps))
  
  ldnow_d<-ldCnow(Xs,R = FALSE,dolog = FALSE, addone=FALSE) %>% upperTmat()
  ldnow<-ldCnow(Xs,R = TRUE,dolog = FALSE, addone=FALSE) %>% upperTmat()
  ldnext<-ldCnext(Xs,w,R=TRUE,dolog = FALSE,addone=FALSE) %>% upperTmat()
  ldnext/ldnow
  Dwith3rd<-append(Dwith3rd,mean(ldnow_d[-1],na.rm=TRUE))
  ldfocal<-append(ldfocal,mean((ldnext/ldnow)[1],na.rm=TRUE))
  
}

qplot(ldfocal,xlab='R between two focal SNPs',xlim=c(0,2))+
  geom_vline(xintercept = mean(ldfocal,na.rm = T), col='firebrick', lty='dashed',lwd=2)
# hist(ldfocal)
mean(ldfocal,na.rm = T)




# It does not change with 2 SNPs
Xs<-Xs[,1:2]
s<-s[1:2]
w<- apply(1+ (s*Xs) , 1, function(i){prod(i)})
ldnow<-ldCnow(Xs,R = TRUE,dolog = FALSE, addone=FALSE) %>% upperTmat()
ldnext<-ldCnext(Xs,w,R=TRUE,dolog = FALSE,addone=FALSE) %>% upperTmat()
ldnext/ldnow

Xs<-Xs[,1:2]
s<-s[1:2]
e<- 0.1
e<- 0
haps<-apply(Xs,1,function(i) prod(i))
w<- apply(1+ s*Xs, 1, function(i){prod(i)})
w<-w^(1+(e*haps))
ldnow<-ldCnow(Xs,R = TRUE,dolog = FALSE, addone=FALSE) %>% upperTmat()
ldnext<-ldCnext(Xs,w,R=TRUE,dolog = FALSE,addone=FALSE) %>% upperTmat()
ldnext/ldnow


# qplot(ldfocal,xlab='R between two focal SNPs')+
#   geom_vline(xintercept = mean(ldfocal,na.rm = T), col='firebrick', lty='dashed',lwd=2)



```




<!-- ```{r} -->



<!-- sa= 0.25 -->
<!-- sb=0.7 -->

<!-- pa=0.5 -->
<!-- pb=0.5 -->

<!-- Dab=0.05 -->


<!-- genos<-expand.grid(c('A','a'),c('B','b')) -->
<!-- genos$g<-paste(genos$Var1,genos$Var2,genos$Var3,sep='') -->
<!-- genos<-data.frame(genos) -->

<!-- #---------------C-------------A-----B------------ -->
<!-- #---------------|-------------|-----|------------ -->

<!-- genos$x<-1 -->
<!-- genos$w<-1 -->

<!-- genos[which(genos$g =='AB'),'w']<- (1+sa)*(1+sb) -->
<!-- genos[which(genos$g =='aB'),'w']<- (1+sb) -->
<!-- genos[which(genos$g =='Ab'),'w']<- (1+sa) -->
<!-- genos[which(genos$g =='ab'),'w']<- 1 -->

<!-- # Dab=Dab/2 -->
<!-- # Dac=Dac/2 -->
<!-- # Dbc=Dbc/2 -->

<!-- genos[which(genos$g =='AB'),'x']<- pa*pb + Dab -->
<!-- genos[which(genos$g =='aB'),'x']<- (1-pa)*pb - Dab -->
<!-- genos[which(genos$g =='Ab'),'x']<- pa*(1-pb) - Dab -->
<!-- genos[which(genos$g =='ab'),'x']<- (1-pa)*(1-pb) + Dab -->
<!-- genos -->

<!-- genos$xprime <- genos$x * genos$w -->
<!-- genos$xprime<-genos$xprime/sum(genos$xprime) -->

<!-- sum(genos$x) -->
<!-- sum(genos$xprime) -->

<!-- pa1= sum(dplyr::filter(genos,Var1 =='A')$xprime)  -->
<!-- pb1 =sum(dplyr::filter(genos,Var2 =='B')$xprime)  -->

<!-- (genos[1,'x'] * genos[4,'x']) - (genos[3,'x'] * genos[2,'x']) -->
<!-- (genos[1,'xprime'] * genos[4,'xprime']) - (genos[3,'xprime'] * genos[2,'xprime']) -->

<!-- (genos[1,'x'] * genos[4,'x']) / (genos[3,'x'] * genos[2,'x']) -->
<!-- (genos[1,'xprime'] * genos[4,'xprime']) / (genos[3,'xprime'] * genos[2,'xprime']) -->
<!-- (genos[1,'w'] * genos[4,'w']) / (genos[3,'w'] * genos[2,'w']) -->


<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='C' & Var2 =='B')$x) * sum(dplyr::filter(genos,Var1 =='c' & Var2 =='b')$x) -->
<!-- )-( -->
<!--   sum(dplyr::filter(genos,Var1 =='C' & Var2 =='b')$x) * sum(dplyr::filter(genos,Var1 =='c' & Var2 =='B')$x) -->
<!-- ) -->
<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$x) -->
<!-- )-( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$x) -->
<!-- ) -->
<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$xprime) -->
<!-- )-( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$xprime) -->
<!-- ) -->

<!-- # D change -->
<!-- ( -->
<!-- ( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$xprime))-( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$xprime))  -->
<!-- )-( -->
<!--   ( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$x))-( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$x) -->
<!-- ) -->
<!-- ) -->

<!-- # r2 change -->
<!-- (( -->
<!--   (( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$xprime))-( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$xprime))  -->
<!--   )/(sqrt(pa*pb*(1-pa)*(1-pb))) -->
<!-- ))-(( -->
<!--   (( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$x))-( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$x) -->
<!--   ))/(sqrt(pa1*pb1*(1-pa1)*(1-pb1))) -->
<!-- )) -->

<!-- ``` -->



<!-- ```{r} -->


<!-- sa= 0.25 -->
<!-- sb=0.7 -->
<!-- sc=0 -->

<!-- pa=0.5 -->
<!-- pb=0.5 -->
<!-- pc=0 -->

<!-- Dab=0.05 -->
<!-- Dac=0 -->
<!-- Dbc=0 -->


<!-- genos<-expand.grid(c('A','a'),c('B','b'),c('C','c')) -->
<!-- genos$g<-paste(genos$Var1,genos$Var2,genos$Var3,sep='') -->
<!-- genos<-data.frame(genos) -->

<!-- #---------------C-------------A-----B------------ -->
<!-- #---------------|-------------|-----|------------ -->

<!-- genos$x<-1 -->
<!-- genos$w<-1 -->

<!-- genos[which(genos$g =='ABC'),'w']<- (1+sa)*(1+sb)*(1+sc) -->
<!-- genos[which(genos$g =='aBC'),'w']<- (1+sb)*(1+sc) -->
<!-- genos[which(genos$g =='AbC'),'w']<- (1+sa)*(1+sc) -->
<!-- genos[which(genos$g =='abC'),'w']<- (1+sc) -->
<!-- genos[which(genos$g =='ABc'),'w']<- (1+sa)*(1+sb) -->
<!-- genos[which(genos$g =='aBc'),'w']<- (1+sb) -->
<!-- genos[which(genos$g =='Abc'),'w']<- (1+sa) -->
<!-- genos[which(genos$g =='abc'),'w']<- 1 -->


<!-- Dab=Dab/2 -->
<!-- Dac=Dac/2 -->
<!-- Dbc=Dbc/2 -->

<!-- genos[which(genos$g =='ABC'),'x']<- pa*pb*pc + Dab+Dac+Dbc -->
<!-- genos[which(genos$g =='aBC'),'x']<- (1-pa)*pb*pc - Dab-Dac+Dbc -->
<!-- genos[which(genos$g =='AbC'),'x']<- pa*(1-pb)*pc - Dab+Dac-Dbc -->
<!-- genos[which(genos$g =='abC'),'x']<- (1-pa)*(1-pb)*pc + Dab-Dac-Dbc -->
<!-- genos[which(genos$g =='ABc'),'x']<- pa*pb*(1-pc) + Dab-Dac-Dbc -->
<!-- genos[which(genos$g =='aBc'),'x']<- (1-pa)*pb*(1-pc) -Dab+Dac-Dbc -->
<!-- genos[which(genos$g =='Abc'),'x']<- pa*(1-pb)*(1-pc) -Dab-Dac+Dbc -->
<!-- genos[which(genos$g =='abc'),'x']<- (1-pa)*(1-pb)*(1-pc) +Dab+Dac+Dbc -->
<!-- genos -->

<!-- genos$xprime<-(genos$w *genos$x)  -->
<!-- genos$xprime<-genos$xprime/ sum(genos$xprime) -->

<!-- sum(genos$x) -->
<!-- sum(genos$xprime) -->

<!-- pa1= sum(dplyr::filter(genos,Var1 =='A')$xprime)  -->
<!-- pb1 =sum(dplyr::filter(genos,Var2 =='B')$xprime)  -->
<!-- pc1 =sum(dplyr::filter(genos,Var3 =='C')$xprime)  -->


<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='C')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='c')$x) -->
<!-- )-( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='c')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='C')$x) -->
<!-- ) -->
<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='C' & Var2 =='B')$x) * sum(dplyr::filter(genos,Var1 =='c' & Var2 =='b')$x) -->
<!-- )-( -->
<!--   sum(dplyr::filter(genos,Var1 =='C' & Var2 =='b')$x) * sum(dplyr::filter(genos,Var1 =='c' & Var2 =='B')$x) -->
<!-- ) -->
<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$x) -->
<!-- )-( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$x) -->
<!-- ) -->
<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$xprime) -->
<!-- )-( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$xprime) -->
<!-- ) -->
<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$xprime) -->
<!-- )-( -->
<!--   pa1 * pb1 -->
<!-- ) -->


<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$x) -->
<!-- )/( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$x) -->
<!-- ) -->
<!-- ( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$xprime) -->
<!-- )/( -->
<!--   sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$xprime) -->
<!-- ) -->


<!-- # # r2 change -->
<!-- # (( -->
<!-- #   (( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$xprime))-( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$xprime) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$xprime))  -->
<!-- #   )/(sqrt(pa*pb*(1-pa)*(1-pb))) -->
<!-- # ))-(( -->
<!-- #   (( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='B')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='b')$x))-( sum(dplyr::filter(genos,Var1 =='A' & Var2 =='b')$x) * sum(dplyr::filter(genos,Var1 =='a' & Var2 =='B')$x) -->
<!-- #   ))/(sqrt(pa1*pb1*(1-pa1)*(1-pb1))) -->
<!-- # )) -->


<!-- ``` -->

<!-- ```{r} -->
<!-- library(dplyr) -->

<!-- ld3loci(sa=0.2,sb=0.2,sc=0,pa=0.5,pb=0.5,pc=0,Dab=0.05,Dac=0,Dbc=0, whatreturn='Rchange') -->
<!-- ld3loci(sa= 0.25,sb=0.7,sc=0,pa=0.5,pb=0.5,pc=0,Dab=0.01,Dac=0,Dbc=0, whatreturn='Rchange') -->
<!-- ld3loci(sa= 0.25,sb=0.7,sc=0,pa=0.5,pb=0.5,pc=0,Dab=0.05,Dac=0,Dbc=0, whatreturn='Rchange') -->


<!-- ``` -->

